---
title: "p8105_hw2_mf3593"
author: "Miao Fu"
date: "2023-09-26"
output: html_document
---
```{r}
library(tidyverse)
library(dplyr)
```

```{r}
pols_df=read_csv("data/pols-month.csv")|>
  janitor::clean_names()|>
  separate(mon, into=c("year","month","day"), "-")|>
  mutate(
    month=toupper(month.abb[as.numeric(month)]),
    president=ifelse(prez_gop==1,"gop","dem"))|>
  select(-c(prez_dem,prez_gop,day))
```
```{r}
snp_df=read_csv("data/snp.csv")|>
  janitor::clean_names()|>
  mutate(date=as.Date(date,format=("%m/%d/%Y")))|>
  separate(date, into=c("year","month","day"), "-")|>
  mutate(year=ifelse(as.numeric(year)>=0 & as.numeric(year)<=9,paste0("200",year),ifelse(as.numeric(year)<=15,paste0("20",year),paste0("19",year))),
         month=toupper(month.abb[as.numeric(month)]))|>
  arrange(month,year)|>
  select(year,month,everything())
```
```{r}
unemploy_df=read_csv("data/unemployment.csv")|>
  janitor::clean_names()|>
  pivot_longer(
    jan:dec,
    names_to = "month",
    values_to= "unemployment"
  )|>
  mutate(month=toupper(month), year=as.character(year))
```
```{r}
snp_pols=left_join(pols_df,snp_df,by=c("year","month"))|>
  left_join(unemploy_df,by=c("year","month"))
```

The pols_df dataset contains data record of `r colnames(pols_df)` from year 1947 to 2015. It records number of people in republican and demographic senators, as well as which party the president is from. It has 822 obs of 9 varaibles. The snp_df dataset contains data record of `r colnames(snp_df)` from year 1950 to 2014. It has 787 obs of 4 variables. The unemploy_df contains data record of `r colnames(unemploy_df)` from year 1948 to 2015. It has 816 obs of 3 variables. The snp_pols combine all three datasets according to year and month columns, and it has 822 obs of 12 variables. 

```{r}
library("readxl")
```

```{r}
trash_df=read_excel("data/trashwheel.xlsx",sheet="Mr. Trash Wheel")|>
  janitor::clean_names()|>
  select(-c("x15","x16"))|>
  drop_na(dumpster)|>
  mutate(homes_powered=(weight_tons*500)/30)
```


