---
title: "p8105_hw2_mf3593"
author: "Miao Fu"
date: "2023-09-26"
output: html_document
---

## Problem 1
```{r}
library(tidyverse)
library(dplyr)
```

```{r}
pols_df=read_csv("data/pols-month.csv")|>
  janitor::clean_names()|>
  separate(mon, into=c("year","month","day"), "-")|>
  mutate(
    month=toupper(month.abb[as.numeric(month)]),
    president=ifelse(prez_gop==1,"gop","dem"))|>
  select(-c(prez_dem,prez_gop,day))
```
```{r}
snp_df=read_csv("data/snp.csv")|>
  janitor::clean_names()|>
  mutate(date=as.Date(date,format=("%m/%d/%Y")))|>
  separate(date, into=c("year","month","day"), "-")|>
  mutate(year=ifelse(as.numeric(year)>=0 & as.numeric(year)<=9,paste0("200",year),ifelse(as.numeric(year)<=15,paste0("20",year),paste0("19",year))),
         month=toupper(month.abb[as.numeric(month)]))|>
  arrange(month,year)|>
  select(year,month,everything(),-day)
```
```{r}
unemploy_df=read_csv("data/unemployment.csv")|>
  janitor::clean_names()|>
  pivot_longer(
    jan:dec,
    names_to = "month",
    values_to= "unemployment"
  )|>
  mutate(month=toupper(month), year=as.character(year))
```
```{r}
snp_pols=left_join(pols_df,snp_df,by=c("year","month"))|>
  left_join(unemploy_df,by=c("year","month"))
```

The `pols_df` dataset contains data record of `r colnames(pols_df)` from year `r range(pols_df$year)[1]` to `r range (pols_df$year)[2]`. It records number of people in republican and demographic senators, as well as which party the president is from. It has `r nrow(pols_df)` rows and `r ncol(pols_df)` columns. The `snp_df` dataset contains data record of `r colnames(snp_df)` from year `r range(snp_df$year)[1]` to `r range (snp_df$year)[2]`. It has `r nrow(snp_df)` rows and `r ncol(snp_df)` columns. The `unemploy_df` contains data record of `r colnames(unemploy_df)` from year `r range(unemploy_df$year)[1]` to `r range (unemploy_df$year)[2]`. It has `r nrow(unemploy_df)` rows and `r ncol(unemploy_df)` columns. The `snp_pols` combine all three datasets according to year and month columns, and it has `r nrow(snp_pols)` rows and `r ncol(snp_pols)` columns. In Januarys in or after 1975 in which a democrat was president, the **average unemployment rate was `r filter(snp_pols, month == "JAN", year >= 1975, president == "dem") |> pull(unemployment) |> mean() |> round(2)`**.  The average unemployment rate over the same time period in which a republican was president was `r filter(snp_pols, month == "JAN", year >= 1975, president == "gop") |> pull(unemployment) |> mean() |> round(2)`.

## Problem 2
```{r}
library("readxl")
```

```{r}
trash_df=read_excel("data/trashwheel.xlsx",sheet="Mr. Trash Wheel")|>
  janitor::clean_names()|>
  select(-c("x15","x16"))|>
  drop_na(dumpster)|>
  mutate(homes_powered=(weight_tons*500)/30,
         setname="Mr.Trash",
         year=as.numeric(year))
```
```{r}
professor_df=read_excel("data/trashwheel.xlsx", sheet="Professor Trash Wheel")|>
  janitor::clean_names()|>
  drop_na(dumpster)|>
  mutate(homes_powered=(weight_tons*500)/30,
         setname="Professor")

Gwynnda_df=read_excel("data/trashwheel.xlsx", sheet="Gwynnda Trash Wheel")|>
  janitor::clean_names()|>
  drop_na(dumpster)|>
  mutate(homes_powered=(weight_tons*500)/30,
         setname="Gwynnda")

combined_df=bind_rows(trash_df,professor_df,Gwynnda_df)|>
  select(setname,everything())

wt_prof=pull(professor_df, weight_tons)|>sum()
Gwynnda_cigarett=pull(Gwynnda_df[Gwynnda_df$month=='July' & Gwynnda_df$year==2021,],cigarette_butts)|>
  sum()
```

There are `r nrow(combined_df)` rows and `r ncol(combined_df)` columns in the `combined_df` data, which combines data from all three datasets. It contains information on quantity collected of different types of trash at each dumpster by Mr.Trash_Wheel, Professor Trash Wheel, and Gynnda Trash Wheel. Mr. Trash_wheel dataset has `r nrow(trash_df)` rows and `r ncol(trash_df)` columns. Mr.Trashwheel dataset records information on `r colnames(trash_df)`. Professor and Gwynnda dataset also collect data on the dumpsters and different types of trash there are. They recorded slightly different types of trash compared to Mr.Trashwheel dataset. For example, professor trash wheel doesn't record number of sports_balls. Gwynnda Trash wheel doesn't have records of glass_bottles, grocery_bags,chip_bags, and sports_balls, but it has records of number of plastic bags which the other two datasets don't have. The total weight of trash collected by Professor Trash Wheel is `r wt_prof`. The total number of cigarette butts collected by Gwynnda in July of 2021 is `r Gwynnda_cigarett`.

## Problem 3
```{r}
bl_df=read_csv("data_mci/MCI_baseline.csv",skip=1)|>
  janitor::clean_names()|>
  mutate(
    sex=case_match( 
      sex,
    1 ~ "Male",
    0 ~ "Female"),
    apoe4=case_match(
      apoe4,
    1 ~ "APOE4 carrier",
    0 ~ "APOE4 non-carrier"
    ))

cleaned_bl=bl_df[bl_df$age_at_onset!='.',]
```
The dataset for baseline MCI includes information on subjects' sex, age, years of education, APOE4 carrier status, and age of onset of AD. A total of 483 subjects were recruited for the baseline study. 97 developed MCI. The average baseline age is `r mean(cleaned_bl$current_age)`. `r (nrow(cleaned_bl[cleaned_bl$sex=='Female'&cleaned_bl$apoe4=='APOE4 carrier',])/nrow(cleaned_bl[cleaned_bl$sex=='Female',]))*100`% of women in the baseline study are APOE4 carriers.

```{r}
amyloid_df=read_csv("data_mci/mci_amyloid.csv", skip=1)|>
  janitor::clean_names()|>
  apply(c(1,2),function(x) ifelse(x=="Na", NA, x))|>
  as.data.frame()|>
  mutate(study_id=as.numeric(study_id))|>
  rename(id=study_id)
```
The dataset for biomarker amyloid includes information on subjects time(in years) elapsed since the study baseline to the visit where biomarker amyloid 42/40 ratio was measured. A total of 5 times of records are included. The dataset has `r nrow(amyloid_df)` rows and `r ncol(amyloid_df)` columns. 

```{r}
bl_only=subset(cleaned_bl, !(id %in% amyloid_df$id))
amyloid_only=subset(amyloid_df, !(id %in% cleaned_bl$id))
combined_df_AD=inner_join(cleaned_bl,amyloid_df,by='id')
write.csv(combined_df_AD, file="data_mci/combined_df.csv")
```

There are `r nrow(bl_only` subjects that is in the baseline study only. There are `r nrow(amyloid_only)` observations in the `amyloid_only` data. The combined dataset of subjects records in both dataset has `r nrow(combined_df_AD)` rows and `r ncol(combined_df_AD)` columns. It combines the varaibles recorded in both datasets, including `r colnames(combined_df_AD)`
